Secure Composition \\ of Robust and Optimising Compilers

### 2023-09-28 csf/main.tex
todos

\vspace{-2em}
\begin{gather*}\small
  \begin{aligned}
  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock
  \qquad
  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
  \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
  \mi{(Pre\text{-}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} 
  \\
                                             & 
                                             \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
  \end{aligned}
\end{gather*}
The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
Since the purpose of the model is to define key security properties of interest, such as \gls*{ms} or a stricter variant of cryptographic constant time, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}~\cite{nagarakatte2010cets}.
%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
Lastly, pre-events describe the actual kind of event that happened.
One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
These two events alone are enough to provide a partial description of \gls*{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
To be able to express \gls*{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
\footnote{
	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
}
Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).

@@

\MPin{
	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
	also, yes, if $\tau$ adds we have a problem, if it removes we don't
\MKin{ Optimisations that remove elements from traces are not necessarily sound,
 e.g., remove all deallocations to "optimise" memory management by offloading
 the work to the OS.

@@

So far, this paper assumes a unified trace-model and secure compilers that do not change the behavior.
It is worth noting that in this setting, the theorem may not be applicable.
For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
By \Thmref{def:trace:tmsdef}, this violates $\tmssafe$.
Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
$\cc{\trg{L}}{\obj{L}}$ modifies the behavior and because of this, the theorem is inapplicable.

This section re-introduces secure compiler composition without this limitation. 
To this end, it utilises existing work on secure compilation~\cite{abate2021extacc} that defines Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
This section is concerned about source-level properties, and henceforth uses the existential image of a source-level property $\src{\pi}$ to map $\src{\pi}$ to a target-level property:

@@

\bul{A cross-language relation $\sim$ is well-formed with respect to $\src{\Pi}$} iff \rul{the existential images preserve membership}.
\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Concrete Property $\src{\Pi}$]\label{def:wf:tracerel}
  $$
  \text{\bul{$\wf{\sim}{\src{\Pi}}$}} \isdef \text{\rul{$\forall \src{\pi}\in\src{\Pi}, \tau_\sim(\src{\pi})\in\tilde{\tau}_\sim(\src{\Pi})$}}
  $$
\end{definition}
\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
  $$
  \wfc{\sim}{\src{\class}} \isdef \forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})
  $$
\end{definition}
\begin{lemma}[Well-Formedness Equivalence for Classes]\label{lem:wf:equiv:wfc}
  $\;\;$
  $ \wf{\sim}{\src{\class}} \Leftrightarrow\ \wfc{\sim}{\src{\class}} $\Coqed
\end{lemma}

@@

To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations:
\begin{definition}[Composing Cross-Language Trace Relations]
  Given $\sim_1 : \src{\trace}\to\irl{\trace}\to\text{Prop},\sim_2 : \irl{\trace}\to\trg{\trace}\to\text{Prop}$, then 
  $$  
  \sim_1\bullet\sim_2\ \isdef \lambda \src{\trace}\ \trg{\trace}. \exists \irl{\trace}. \src{\trace}\sim_1\irl{\trace} \text{ and } \irl{\trace}\sim_2\trg{\trace}
  $$
\end{definition}

\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
  If $\rtpsim{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}{\sim_1}$, $\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\class[_{2}]\right)}{\sim_2}$, and $\wf{\sim_1}{\class[_2]}$, then $\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}{\sim_1\bullet\sim_2}$. \Coqed
\end{theorem}

@@

With this, the previous claim follows as a simple corollary:
\begin{corollary}
  $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$.
\end{corollary}
\begin{proof}
  Note that $\neg\left(\wf{\sim_{{\smssafe}}}{{\tmssafe}}\right)$, since the bounds-table is never deallocated and therefore, for any $\trg{\trace}\in\tmssafe$, $\tau_{\sim_{\tmssafe}}\left(\trg{\trace}\right)\notin\tilde{\tau}_{\sim_{\tmssafe}}\left({\tmssafe}\right)$.
  However, by assumption, $\wf{\sim_{{\tmssafe}}}{{\tmssafe}}$.
  Using \Thmref{lem:useless} it follows that $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\right)=\emptyset$.
\end{proof}

>>>>

\vspace{-1em}
\begin{gather*}\small
  \begin{aligned}
  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock
  \qquad
  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
  \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
  \mi{(Pre\text{-}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} 
  \\
                                             & 
                                             \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
  \end{aligned}
\end{gather*}
The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
Since the purpose of the model is to define key security properties of interest, such as \gls*{ms} or a stricter variant of cryptographic constant time, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls*{tms}~\cite{nagarakatte2010cets}.
%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
Lastly, pre-events describe the actual kind of event that happened.
One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
These two events alone are enough to provide a partial description of \gls*{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
To be able to express \gls*{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
Typical operations with data-dependent timing are branches and certain binary operations, such as division.%
\footnote{
	This is architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction~\cite[p.~755]{arm-refman}.
}
Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).

@@

% \MPin{
% 	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
%  e.g., remove all deallocations to "optimise" memory management by offloading
%  the work to the OS.

@@

So far, this paper assumes a unified trace-model and secure compilers that do not change the behavior.
It is worth noting that in this setting, the theorem may not be applicable.
For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
By \Thmref{def:trace:tmsdef}, this violates $\tmssafe$.
Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
$\cc{\trg{L}}{\obj{L}}$ modifies the behavior and because of this, the theorem is inapplicable.
\MPin{
	this is still applicable though
	many use it, and we showcase it too.
}

This section re-introduces secure compiler composition without this limitation. 
To this end, it utilises existing work on secure compilation~\cite{abate2021extacc} that defines Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
This section is concerned about source-level properties, and henceforth 
\MPin{
	different source and target trace models
	.
	traces are related
	.
	there is a function tau that lefts the relation from individual traces to props/hprops
}
\MPin{
	drop next 3 defs
}

uses the existential image of a source-level property $\src{\pi}$ to map $\src{\pi}$ to a target-level property:

@@

\bul{A cross-language relation $\sim$ is well-formed with respect to $\src{\Pi}$} iff \rul{the existential images preserve membership}.
\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Concrete Property $\src{\Pi}$]\label{def:wf:tracerel}
  $$
  \text{\bul{$\wf{\sim}{\src{\Pi}}$}} \isdef \text{\rul{$\forall \src{\pi}\in\src{\Pi}, \tau_\sim(\src{\pi})\in\tilde{\tau}_\sim(\src{\Pi})$}}
  $$
\end{definition}
\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
  $$
  \wfc{\sim}{\src{\class}} \isdef \forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})
  $$
\end{definition}
\MPin{
	explain with text and underline
}
\begin{lemma}[Well-Formedness Equivalence for Classes]\label{lem:wf:equiv:wfc}
  $\;\;$
  $ \wf{\sim}{\src{\class}} \Leftrightarrow\ \wfc{\sim}{\src{\class}} $\Coqed
\end{lemma}

@@

To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations:
\MPin{
	do in text
}
\begin{definition}[Composing Cross-Language Trace Relations]
  Given $\sim_1 : \src{\trace}\to\irl{\trace}\to\text{Prop},\sim_2 : \irl{\trace}\to\trg{\trace}\to\text{Prop}$, then 
  $$  
  \sim_1\bullet\sim_2\ \isdef \lambda \src{\trace}\ \trg{\trace}. \exists \irl{\trace}. \src{\trace}\sim_1\irl{\trace} \text{ and } \irl{\trace}\sim_2\trg{\trace}
  $$
\end{definition}

\MPin{
	explain
}
\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
  If $\rtpsim{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}{\sim_1}$, $\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\class[_{2}]\right)}{\sim_2}$, and $\wf{\sim_1}{\class[_2]}$, then $\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}{\sim_1\bullet\sim_2}$. \Coqed
\end{theorem}

@@

\MPin{
	drop
}
With this, the previous claim follows as a simple corollary:
\begin{corollary}
  $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$.
\end{corollary}
\begin{proof}
  Note that $\neg\left(\wf{\sim_{{\smssafe}}}{{\tmssafe}}\right)$, since the bounds-table is never deallocated and therefore, for any $\trg{\trace}\in\tmssafe$, $\tau_{\sim_{\tmssafe}}\left(\trg{\trace}\right)\notin\tilde{\tau}_{\sim_{\tmssafe}}\left({\tmssafe}\right)$.
  However, by assumption, $\wf{\sim_{{\tmssafe}}}{{\tmssafe}}$.
  Using \Thmref{lem:useless} it follows that $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\right)=\emptyset$.
\end{proof}

\MPin{
	what about sim2 and C1?
	-> no relation 
}
\MPin{
	For the sake of simplicity we go back to the singel trace model and go back to the framework of single trace model
}
